A Simple and Strong Baseline for Universal Targeted Attacks on Siamese Visual Tracking

### 2021-05-01 博士/论文提交/adversarial_attack/TCSVT/latex/imperceptible.tex
改拼写

A straight forward way to achieve targeted attack for Siamese trackers is directly using popular write-box attack methods such as FGSM \cite{FGSM} and BIM \cite{DBLP:conf/iclr/KurakinGB17a}:

\begin{equation}
\begin{gathered}
  \min\limits_{\delta_{\textbf{x}_{i}}, \delta_{\textbf{z}_{i}}} L(y^{fake}_i, f(\textbf{x}_i + \delta_{\textbf{x}_{i}}, \textbf{z}_i + \delta_{\textbf{z}_{i}}))\\
  \text{s.t.}\ ||\delta_{\textbf{x}_i}||_p \le \epsilon_1, ||\delta_{\textbf{z}_i}||_p \le \epsilon_2,
\end{gathered}
\end{equation}
where $L$ is the loss function of the SiamFC++ tracker. $(\textbf{x}_i, \textbf{z}_i)$ is the search-template pair of the video being tracked at frame $i$ and $\textbf{z}_i \equiv \textbf{z}_1$. $y^{fake}_i=\{\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*\}$ is the fake target of the loss function generated according to the fake ground truth box $b^{fake}_i$ that we want the tracker to output at frame $i$ (see Sec. \ref{generate}). 
% 不写 BIM，原因是BIM的优化为 xi+1 = xi + delta, 与上面公式不太好兼容。
Additionally, $\delta$ should be sufficiently small, which is commonly modeled through an upper-bound $\epsilon$ on the $l_p\text{-norm}$, commonly denoted as $||\cdot||_p$, of the perturbation. A popular choice is to set $p=\infty$.
However, these methods require perturbing the input image pair $(\textbf{x}, \textbf{z})$ frame by frame, which comes at a non-negligible cost, considering that VOT is a real-time task.
As a consequence, universal adversarial perturbations \cite{UAP, shafahi2020universal} are more practical for attacking Siamese trackers:
\begin{equation}
  \begin{gathered}
    \min\limits_{\delta_\textbf{x}, \delta_\textbf{z}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z})} L(y^{fake}, f(\textbf{x} + \delta_\textbf{x}, \textbf{z} + \delta_\textbf{z}))\\
    \text{s.t.}\ ||\delta_\textbf{x}||_p \le \epsilon_1, ||\delta_\textbf{z}||_p \le \epsilon_2,
  \end{gathered}
\end{equation}  
where the input pair $(\textbf x,\textbf z)$ is randomly picked from the training set. 
Note that universal adversarial perterbations are added to the whole image and cannot tell the tracker where to look, so $y^{fake}$ is kept constant during the training process, resulting in the inability to perform targted attacks.
To achieve universal targeted attacks, one feasible way is to paste a small (universal) adversarial patch into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the adversarial patch instead of the real target:
\begin{equation}
    \min\limits_{\delta_\textbf{x}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z}, y^{fake})} L(y^{fake}, f(\textbf{x}\odot (1-m) + \delta_\textbf{x}\odot m, \textbf{z})),
\end{equation}
where $m$ is a binary mask and $m_{x,y}=1$ at all spatial locations under the fake ground bounding box $b^{fake}$ which is variable during training according to the patch position. $\odot$ is element-wise multiplication. Note that this vanilla adversarial patch method uses the clean template image during training. However, CNN attacks are usually expected imperceptible but the above method has to paste an obviously noticeable fake target patch to tracking frames, which raises the risk of being suspected.

>>>>

A straight forward way to achieve targeted attacks for Siamese trackers is directly using popular write-box attack methods such as FGSM \cite{FGSM} and BIM \cite{DBLP:conf/iclr/KurakinGB17a}:

\begin{equation}
\begin{gathered}
  \min\limits_{\delta_{\textbf{x}_{i}}, \delta_{\textbf{z}_{i}}} L(y^{fake}_i, f(\textbf{x}_i + \delta_{\textbf{x}_{i}}, \textbf{z}_i + \delta_{\textbf{z}_{i}}))\\
  \text{s.t.}\ ||\delta_{\textbf{x}_i}||_p \le \epsilon_1, ||\delta_{\textbf{z}_i}||_p \le \epsilon_2,
\end{gathered}
\end{equation}
where $L$ is the loss function of the SiamFC++ tracker. $(\textbf{x}_i, \textbf{z}_i)$ is the search-template pair of the video being tracked at frame $i$ and $\textbf{z}_i \equiv \textbf{z}_1$. $y^{fake}_i=\{\textbf{C}^*, \textbf{R}^*, \textbf{Q}^*\}$ is the fake target of the loss function generated according to the fake ground truth box $b^{fake}_i$ that we want the tracker to output at frame $i$ (see Sec. \ref{generate}). 
% 不写 BIM，原因是BIM的优化为 xi+1 = xi + delta, 与上面公式不太好兼容。
Additionally, $\delta$ should be sufficiently small, which is commonly modeled through an upper-bound $\epsilon$ on the $l_p\text{-norm}$, commonly denoted as $||\cdot||_p$, of the perturbation. A popular choice is to set $p=\infty$.
However, these methods require perturbing the input image pair $(\textbf{x}, \textbf{z})$ frame by frame, which comes at a non-negligible cost, considering that VOT is a real-time task.
As a consequence, universal adversarial perturbations \cite{UAP, shafahi2020universal} are more practical for attacking Siamese trackers:
\begin{equation}
  \begin{gathered}
    \min\limits_{\delta_\textbf{x}, \delta_\textbf{z}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z})} L(y^{fake}, f(\textbf{x} + \delta_\textbf{x}, \textbf{z} + \delta_\textbf{z}))\\
    \text{s.t.}\ ||\delta_\textbf{x}||_p \le \epsilon_1, ||\delta_\textbf{z}||_p \le \epsilon_2,
  \end{gathered}
\end{equation}  
where the input pair $(\textbf x,\textbf z)$ is randomly picked from the training set. 
Note that universal adversarial perterbations are added to the whole image and cannot tell the tracker where to look, so $y^{fake}$ is kept constant during the training process, resulting in the inability to perform targeted attacks.
To achieve universal targeted attacks, one feasible way is to paste a small (universal) adversarial patch into the search images adhering to the predefined trajectory, so that the tracker outputs the location and size of the adversarial patch instead of the real target:
\begin{equation}
    \min\limits_{\delta_\textbf{x}} \mathop{\mathbb{E}}\limits_{(\textbf{x}, \textbf{z}, y^{fake})} L(y^{fake}, f(\textbf{x}\odot (1-m) + \delta_\textbf{x}\odot m, \textbf{z})),
\end{equation}
where $m$ is a binary mask and $m_{x,y}=1$ at all spatial locations under the fake ground bounding box $b^{fake}$ which is variable during training according to the patch position. $\odot$ is element-wise multiplication. Note that this vanilla adversarial patch method uses the clean template image during training. However, CNN attacks are usually expected imperceptible but the above method has to paste an obviously noticeable fake target patch to tracking frames, which raises the risk of being suspected.
